Warming Degrees and Flowering Trees

### 2025-04-04 paper/paper.qmd
Revise README to enhance project overview and clarify data structure for sakura flowering analysis

The regression analysis in @tbl-reg-results reveals a significant inverse relationship between March temperature and cherry blossom blooming dates. For each 1°C increase in March temperature, cherry blossoms bloom approximately 2.25 days earlier (p < 0.01). This relationship remains consistent across both standard and robust OLS models, though the latter shows larger standard errors, suggesting heteroscedasticity in the data. With an R-squared of 0.32, March temperature appears to explain about one-third of the variation in blooming dates. When March temperature is 0°C, blossoms are predicted to emerge on the 88th day of the year (late March). These findings indicate that spring temperature serves as a meaningful predictor of cherry blossom phenology.

```{python}
#| label: fig-reg-plt 
#| fig-cap: "Regression Visualization"

@@

```{python}
# Scatterplot with year-based color gradient
plt.figure(figsize=(10, 6))
scatter = plt.scatter(df['march_temp_c'], df['flower_doy'], c=df['year'], cmap='viridis', s=60, edgecolor='k')
plt.plot(X['march_temp_c'], model_0.predict(X), color='red', label='OLS Fit')
plt.text(-0.2, 95, f"blooming_day = {model_0.params[1]:.2f} * temp + {model_0.params[0]:.2f}", color="red", fontsize=12)
plt.colorbar(scatter, label='Year')
plt.title("Sakura Blooming DoY vs March Temperature with Year Gradient")
plt.xlabel("March Avg Temperature (°C)")
plt.ylabel("Blooming Day of Year (DoY)")
plt.legend()
plt.show()
```

@@

## Cultural Implications
From a cultural perspective, the temporal shift in cherry blossom seasons need adjustments to traditional hanami celebrations and tourism schedules, which have historically been calibrated to conventional blooming periods. In addition, these modifications could impact cultural practices that have remained relatively consistent for centuries and negatively impact tourism economy that is built around it.

## Limitations
We would like to address several methodological limiting factors. Firstly, our analysis relies predominantly on data from Kyoto, representing a geographically limited subset that may not comprehensively reflect sakura phenological patterns across Japan's diverse climate regions. Secondly, although we used a merged dataset spanning from 800 CE to 2019, there are significant data gaps/missing values where we have to drop during the cleaning & merging of the two dataset. In addition, a notable limitation is the historical dataset's reliability, as large portions comprise of reconstructed values derived from local linear regression models or originated from primary sources that lack rigorous academic verification protocols. This presents challenges for establishing data provenance and reliability in accordance with contemporary scientific standards. Lastly, while our findings highlight temperature as a statistically significant predictor of bloom timing, the single-variable approach inherently overlooks other potential ecological factors. Variables such as precipitation patterns, photoperiod fluctuations, soil conditions, and winter chilling requirements may also contribute significantly to the observed advancement in flowering dates. A more comprehensive multivariate model would potentially yield enhanced predictive capacity and ecological insights.

@@

>>>>

The regression analysis in @tbl-reg-results reveals a significant inverse relationship between March temperature and cherry blossom blooming dates. For each 1°C increase in March temperature, cherry blossoms bloom approximately 2.25 days earlier (p < 0.01). This relationship remains consistent across both standard and robust OLS models, though the latter shows larger standard errors, suggesting heteroscedasticity in the data. With an R-squared of 0.32, March temperature appears to explain about one-third of the variation in blooming dates. When March temperature is 0°C, blossoms are predicted to emerge on the 88th day of the year (late March). These findings indicate that spring temperature serves as a meaningful predictor of cherry blossom phenology. Diagnostic plots are in @sec-appendix-1

```{python}
#| label: fig-reg-plt 
#| fig-cap: "Regression Visualization"
#| message: false
#| echo: false
#| warning: false
#| fig-align: center

@@

```{python}
#| label: fig-train-plt
#| fig-cap: "Trained Predictions Visualization"
#| message: false
#| echo: false
#| warning: false
#| fig-align: center

# Test OLS model (nonrobust)

model_test = model_0

# Prepare test data
X_test = sm.add_constant(test_df['march_temp_c'])
y_test = test_df['flower_doy']

# Predict on test data
y_pred = model_test.predict(X_test)

# Plot actual vs predicted
plt.figure(figsize=(10, 6))
plt.plot(test_df['year'], y_test, label='Actual Blooming DoY', marker='o')
plt.plot(test_df['year'], y_pred, label='Predicted Blooming DoY', marker='x')
plt.xlabel("Year")
plt.ylabel("Blooming Day of Year")
plt.title("Actual vs Predicted Sakura Blooming (2001–2019)")
plt.legend()
plt.grid(True)
```

@@

## Cultural Implications
As established in @sec-lit-review, that tradition of Hanami is a staple in Japanese culture. Moreover, it has become a significant source of revenue for both the nation itself and the local communities which host associated festivals. Our findings reveal a significant relationship between pre-flowering temperatures and blooming dates, aligning closely with observed patterns with climate-change related temperature increases. As the Hanami/cherry blossom festivals draw significant international and domestic attention, the scheduling of these events must be exhaustive in order to facilitate them at the scale they reach (and continue to grow). While we don’t anticipate a significant disruption to the bloom itself, understanding when the bloom will start is critical to maintaining the economic momentum these festivals have demonstrated. Furthermore, cherry blossoms and the Hanami practice are significant beyond their economic benefits. Communities and citizens celebrate and find great meaning and importance in these practices. Understanding that the bloom may come earlier, and being able to plan for travel, accommodations, and individual scheduling is important to many people in Japan. We expect that, if temperatures continue to raise year over year, it may disrupt the lives/scheduling of communities in Japan which find great importance in observing cherry blossom blooming every year.

## Limitations
There are several limitations to this analysis. First, our analysis relies predominantly on data from Kyoto, representing a geographically limited subset that may not comprehensively reflect sakura phenological patterns across Japan's diverse climate regions. Second, although we used a merged dataset spanning from 800 CE to 2019, there are significant data gaps/missing values where we have to drop during the cleaning & merging of the two dataset. In addition, a notable limitation is the historical dataset's reliability, as large portions comprise of reconstructed values derived from local linear regression models or originated from primary sources that lack rigorous academic verification protocols. This presents challenges for establishing data provenance and reliability in accordance with contemporary scientific standards. Lastly, while our findings highlight temperature as a statistically significant predictor of bloom timing, the single-variable approach inherently overlooks other potential ecological factors. Variables such as precipitation patterns, photoperiod fluctuations, soil conditions, and winter chilling requirements may also contribute significantly to the observed advancement in flowering dates. A more comprehensive multivariate model, comprising more reliable data would potentially yield enhanced predictive capacity and ecological insights.

@@

## Diagnostic Plots for OLS Models {#sec-appendix-1}

```{python}
#| fig-cap: "Diagnostic Results for Cherry Blossom Flowering Day vs. March Temperature"
#| message: false
#| echo: false
#| warning: false
#| fig-align: center

# Add necessary libraries for diagnostic plots
import statsmodels.api as sm
from statsmodels.graphics.gofplots import qqplot
import matplotlib.pyplot as plt
import seaborn as sns
import numpy as np

# Create diagnostic plots for the OLS model
fig = plt.figure(figsize=(20, 15))

# Residual vs Fitted Plot
ax = fig.add_subplot(1, 1, 1)
residuals = model_0.resid
fitted = model_0.fittedvalues
sns.residplot(x=fitted, y=residuals, lowess=True, scatter_kws={'alpha': 0.5}, line_kws={'color': 'red', 'lw': 1}, ax=ax)
ax.set_title('Residuals vs Fitted')
ax.set_xlabel('Fitted values')
ax.set_ylabel('Residuals')
ax.axhline(y=0, color='k', linestyle='-', alpha=0.3)
```

```{python}
#| message: false
#| echo: false
#| warning: false
#| fig-align: center

import statsmodels.api as sm
from statsmodels.graphics.gofplots import qqplot
import matplotlib.pyplot as plt
import seaborn as sns
import numpy as np

# Create diagnostic plots for the OLS model
fig = plt.figure(figsize=(10, 8))
ax = plt.subplot(111)

# QQ Plot for Normality
qqplot(model_0.resid, line='s', ax=ax)
ax.set_title('Normal Q-Q')
ax.set_xlabel('Theoretical Quantiles')
ax.set_ylabel('Standardized Residuals')
```

```{python}
#| message: false
#| echo: false
#| warning: false
#| fig-align: center

# Add necessary libraries for diagnostic plots
import statsmodels.api as sm
from statsmodels.graphics.gofplots import qqplot
import matplotlib.pyplot as plt
import seaborn as sns
import numpy as np

# Create diagnostic plots for the OLS model
fig = plt.figure(figsize=(10, 8))

# Scale-Location Plot (Square root of absolute residuals vs fitted)
ax = fig.add_subplot(2, 2, 3)
sqrt_abs_resid = np.sqrt(np.abs(residuals))
sns.regplot(x=fitted, y=sqrt_abs_resid, lowess=True, scatter_kws={'alpha': 0.5}, line_kws={'color': 'red', 'lw': 1}, ax=ax)
ax.set_title('Scale-Location')
ax.set_xlabel('Fitted values')
ax.set_ylabel('$\\sqrt{|Standardized Residuals|}$')

```{python}
#| message: false
#| echo: false
#| warning: false
#| fig-align: center

# Add necessary libraries for diagnostic plots
import statsmodels.api as sm
from statsmodels.graphics.gofplots import qqplot
import matplotlib.pyplot as plt
import seaborn as sns
import numpy as np

# Leverage vs. Standardized Residuals
fig = plt.figure(figsize=(10, 8))
ax = fig.add_subplot(1, 1, 1)
infl = model_0.get_influence()
hatvals = infl.hat_matrix_diag
stud_resid = infl.resid_studentized_external

# Determine leverage threshold
p = len(model_0.params)  # number of parameters
leverage_threshold = 2 * p / len(hatvals)

# Plot points
ax.scatter(hatvals, stud_resid, alpha=0.6)
ax.axhline(y=0, color='k', linestyle='-', alpha=0.3)

# Add reference lines
ax.axhline(y=2, color='g', linestyle='--', label='Std. Residuals = 2')
ax.axhline(y=-2, color='g', linestyle='--')
ax.axvline(x=leverage_threshold, color='b', linestyle='--', 
           label=f'Leverage threshold ({leverage_threshold:.3f})')

# Highlight and label influential points (leverage > threshold or std. residuals > 2)
influential_points = []
for i in range(len(hatvals)):
    if (abs(stud_resid[i]) > 2) or (hatvals[i] > leverage_threshold):
        ax.text(hatvals[i], stud_resid[i], 
                f" {i} ({hatvals[i]:.3f}, {stud_resid[i]:.2f})", 
                fontsize=9, color="red")
        influential_points.append(i)

ax.set_title('Leverage vs. Standardized Residuals')
ax.set_xlabel('Leverage (hat values)')
ax.set_ylabel('Studentized Residuals')
ax.legend(loc='upper right', fontsize=8)
plt.tight_layout()
```
